---
title: "Replication Report on Enos (2014)"
author: "David Kane"
output:
  html_document
citation_package: natbib
bibliography: bibliography.bib
---


## Abstract

I replicate "Causal effect of intergroup contact on exclusionary attitudes" (@enos2014) using code/data deposited in the Harvard Dataverse (@DVN/DOP4UB_2017). There are two figures and three tables in the original article. Figure 2 and Tables 1 and 2 can be, almost, perfectly replicated. Table 3 and Figure 1 can not be replicated from the code/data on deposit at the Dataverse.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(ri)
library(RItools)
library(xtable)
library(gt)
library(tidyverse)

# The raw data does not have a column name for the first column, so, by default
# it gets X1. But this looks to be just a row number, so let's ignore. 

# The variable names in dat_all are just pathetic, as if they are the result of
# an extremely sloppy merge. What is the difference, for example, between
# party.x and party.y?

# Using read_csv() causes lots of problems later one because dat.all, for
# example, as a tibble behaves differently than dat.all as a data frame. The key
# issue is that when you select out a single column, a data frame will
# automatically drop dimensions and become a vector. A tibble requires you to
# specify drop = TRUE.

dat.all <- read_csv("data/pnas_data.csv") %>% 
  
  # Instead of keeping around 90 variables --- many of them garbage? --- let's
  # just keep the variables we need. Most of these are for Table 2.
  
  select(habits, station, treated_unit,
         numberim.x, Remain.x, Englishlan.x,
         numberim.y, Remain.y, Englishlan.y, 
         line.x, t.time,
         treatment, liberal, republican, obama.disapprove, ride.everyday, 
         voted.2010, romney.voter, Hispanics.x, age, residency.new, college, 
         male, hispanic.new,  white, income.new)

# There are several other csv files but, as best I can tell, they are only
# relevant to the supplementary materials.

# Always nice to do some checks which confirm claims in the article, although
# note that, in the analysis, one of the stations is dropped because of a lack
# of data on the control arm.

stopifnot(length(table(dat.all$station)) == 9)

```

## Table 1. Experiment results

I am able to replicate all the numbers from Table 1. One clarification is that the values for *n* are somewhat misleading. *n* varies across the questions. For example, *n* for the question about "English as official language?" is 109, as stated. But the *n* for "Number of immigrants be increased" is only 106. It is, obviously, sloppy to list a single *n* when, in fact, *n* varies across questions. The same issue arises with regard to the subset of commuters who wait in the platform. The *n* for "Number of immigrants be increased" is only 98, as compared to the stated 100 for the other two questions. This is a minor point with no meaningful implication for the results.

```{r table_1}
# Note how hard it is to grok this code without comments. And, even if you wrote
# it yourself, it would all be opaque two years from now without comments.

# There are three key questions and, for some reason, .x and .y versions of
# those variables. Why?

repeats <- c("numberim", "Remain", "Englishlan")

x.names <- paste(repeats, ".x", sep = "")
y.names <- paste(repeats, ".y", sep = "")

# Which train line is definately important. What would happened if we handled this differently?

covariates <- c('line.x')

# final.mat is where we are going to place the results as we loop through.

final.mat <- matrix(nrow = 0, ncol = 8)

# We need to do the same code for both all individuals and for those who wait on
# the platform (hence "no.car"). By the way, how did they figure out ahead of
# time about the importance of waiting in cars?

subsets <- c('all', 'no.car')

# This code is absurd. Not the least is using "subset" --- the name of a base
# function! --- as a character vector. Worth the hour (or more!) it would take
# to refactor?

for(subset in subsets){

	out.mat = matrix(nrow = length(repeats), ncol = 8)
	
	if(subset == 'all'){
		dat.subset = dat.all
		}
	if(subset ==  'no.car'){
		dat.subset = dat.all[dat.all$habits != 1,]
		}

		
	z.variable = 'treatment'
	
	# This portion is really tricky because x.new and friends gets created as a
	# data.frame, which is now just an element of the dat.subset list. That is
	# weird enough, but then, later, we try to work with it and bad stuff happens.
	# I suspect that all this is fine if we just never turned the initial data into
	# a tibble. But maybe not! And, even if it does work, it is a bad idea. Trick
	# was to add "drop = TRUE", which ensures that the resulting object is a
	# vector.
	
	for(j in 1:length(repeats)){
		dat.subset$x.new <- (dat.subset[, x.names[j], drop = TRUE]- 1)/4  ##rescale x to 0-1
		dat.subset$y.new <- (dat.subset[, y.names[j], drop = TRUE]- 1)/4  ##rescale y to 0-1
		dat.subset$Y     <-  dat.subset$y.new - dat.subset$x.new
		
		dat.use <- dat.subset[is.na(dat.subset$Y) == F, ]
		
		x.sd = sd(as.vector(dat.use$x.new),na.rm = T)
		x.mean = mean(dat.use$x.new,na.rm = T)
		y.mean = mean(dat.use$y.new,na.rm = T)
		y.treat = mean(dat.use$y.new[dat.use$treatment==1],na.rm = T)
	
		station.treatment.table <- table(dat.use$station, dat.use[ ,z.variable, drop = TRUE])
		no.control.stations = names(which(station.treatment.table[,1] == 0))
		no.treatment.stations = names(which(station.treatment.table[,2] == 0))
		dat.use = dat.use[!dat.use$station%in%c(no.control.stations,no.treatment.stations),]
				
		
		dat.use$station = factor(dat.use$station)
		dat.use$treated_unit = factor(dat.use$treated_unit)
		Xs = data.matrix(dat.use[,covariates])
		
		perms <- genperms(Z = dat.use[,z.variable], blockvar=dat.use$station, clustvar=dat.use$treated_unit)
		probs = genprobexact(Z = dat.use[,z.variable], blockvar=dat.use$station, clustvar=dat.use$treated_unit)

		ate = estate(Y = dat.use$Y, Z = dat.use[, z.variable, drop = TRUE], X = Xs, prob = probs)
		Ys = genouts(Y = dat.use$Y, Z = dat.use[, z.variable, drop = TRUE], ate = 0)
		distout <- gendist(Ys,perms, prob=probs)
		disp =	dispdist(distout, ate = ate, display.plot = F)
		
		out.mat[j,1] = repeats[j]
		out.mat[j,2] = subset
		out.mat[j,3] = nrow(dat.use)
		out.mat[j,4] = ate
		out.mat[j,5] = disp$greater.p.value
		out.mat[j,6] = disp$lesser.p.value
		out.mat[j,7] = x.sd
		out.mat[j,8] = x.mean
		}
	final.mat = rbind(final.mat, out.mat)
}

# Clean up the result and save it for later.

final.mat <- as.data.frame(final.mat)
colnames(final.mat) <- c('variable','subset','N','ate','greater.p.value','lesser.p.value','x.sd','x.mean')


final.mat.main <- final.mat 

```



```{r output_create, results="asis"}
# OMG! He calculates many (most?) of these numbers as factors! Check out
# sapply(final.mat.main, class). (Or is that somehow my fault for using tibbles
# instead of data frames?) Then he has to convert to character and then to
# numeric to use them properly. Insanity! 

# But, also, there should be a better way to clean this up.

final.mat.main$ate    <- as.numeric(as.character(final.mat.main$ate)) 
final.mat.main$N      <- as.numeric(as.character(final.mat.main$N)) 
final.mat.main$x.mean <- as.numeric(as.character(final.mat.main$x.mean)) 
final.mat.main$x.sd   <- as.numeric(as.character(final.mat.main$x.sd)) 
final.mat.main$greater.p.value <- as.numeric(as.character(final.mat.main$greater.p.value)) 

# Now that we have final.mat.main, we can make some nice output. The annoying
# part is that this data is very poorly organized, if our main purpose is to
# create Table 1. I really ought to go back and recalculate these data frames in
# a tidyier fashion. But, for now, we will just pick out the parts we want.

# This was annoyingly hard, which goes to show why it is nice to calculate
# important numbers in a sensible way in the first place. Main trick is to
# create something of a convenient size, fill in the numbers and then start
# tabling. Must be easier ways to do this.

as_tibble(x = matrix(data = NA_real_, 
                     nrow = 4, 
                     ncol = 6,
                     dimnames = list(NULL, paste0("V", 1:6)))) %>%  
  
  # Remember: tibbles can't have row names anymore. So, we need to make them a
  # variable.
  
  mutate(question = c('Number of immigrants be increased?',
                      'Children of undocumented be allowed to stay?',
                      'English as official language?', 
                      'n')) %>% 
  mutate(V1 = c(final.mat.main$ate[1:3], final.mat.main$N[2]),
         V2 = c(final.mat.main$ate[4:6], final.mat.main$N[5]),
         V3 = c(final.mat.main$x.mean[1:3], final.mat.main$N[2]),
         V4 = c(final.mat.main$greater.p.value[1:3], NA),
         V5 = c(final.mat.main$greater.p.value[4:6], NA),
         V6 = c(final.mat.main$x.sd[1:3], NA)) %>% 

  #  Start charting!

  gt() %>% 
  tab_header(title = "Table 1. Experiment results") %>% 
  
  cols_move_to_start(columns = vars(question)) %>% 
  
  fmt_number(columns = starts_with("V"), decimals = 3, drop_trailing_zeros = TRUE) %>%
  
  # Combine the columns.
  
  cols_merge(col_1 = vars(V1), col_2 = vars(V4), pattern = "{1} ({2})") %>% 
  cols_merge(col_1 = vars(V2), col_2 = vars(V5), pattern = "{1} ({2})") %>% 
  cols_merge(col_1 = vars(V3), col_2 = vars(V6), pattern = "{1} ({2})") %>% 
  
  cols_label(question =  "Question",        
             V1 = "ATE (P)",
             V2 = "CATE (P)",
             V3 = "T1 levels (SD)") %>% 
  
  cols_move(columns = vars(V3), after = vars(question)) %>% 
  
  tab_spanner(label = "All respondents",   columns = vars(V1, V3)) %>% 
  tab_spanner(label = "Waits on platform", columns = vars(V2)) %>% 

  tab_source_note(source_note = 'In the first "All respondents" column, ATE represents responses in T2-T1 for the treatment group compared with the control group for the entire experimental sample. Positive values mean a more politically conservative response. In the "Waits on platform" column, CATEs are the Conditional Average Treatment Effects for persons who said they stand on the platform, rather than wait in their cars. In the second "All respondents" column, T1 levls and SDs for each variable for all respondents. All variables are scaled 0--1.') %>% 
  
  tab_options(footnote.glyph = c("*", "&dagger;")) %>%   
  
  tab_footnote(
    footnote = "P values from a one-tailed test against the Null Hypothesis of no effect are in parantheses.",
    locations = cells_column_labels(
      columns = vars(V1))) %>% 
  
  tab_footnote(
    footnote = "Each of the questions allowed responses on a five-point scale ranging from strongly agree to stringly disagree (exact answers were changed to be appropriate to the actual question.",
    locations = cells_data(
      columns = vars(question),
      rows = 1)) %>% 

  # Should be possible to replace as_raw_html() with as_latex() and then knit to
  # PDF. Alas, I can't get it to work. I suspect that gt may be doing some
  # tricky latex code which the default packages has trouble with. Ought to try
  # to simplify this to a reproducible example, even dive into the .tex output.
  
  as_raw_html() %>% as.character() %>% cat()

# Discuss the confusion about exactly what those numbers mean. Does a high
# number mean the conservative answer?


```

\newpage

## Table 2

I am able to replicate all the values for Table 2, except for those related to income. But, even there, my results are very close to those published.

```{r table_2, results="asis", tab.cap = NULL}
# I believe that this is the code which drives Table 2. I deleted the other
# tests. Note that income is present in the published table but at a different
# location than in the code. I could not get replicate the income numbers.
# Error?

out.balance.test <- xBalance(fmla = treatment ~ liberal + republican + obama.disapprove + 
                               ride.everyday + voted.2010 + romney.voter + Hispanics.x + 
                               age + residency.new + college + male +
                               hispanic.new +  white + income.new, data = dat.all, 
                             report = c("std.diffs","z.scores","adj.means"), 
                             strata = factor(dat.all$station))

# This creates an object of class xbal, which is obviously too annoying to work
# with since it is also a list with two items, and the first item is very
# strange. However, the below incantation gets out the matrix of numbers we care
# about. Is there a better way?

# Once I have the data, I add a labelling var before getting to gt.

x <- as_tibble(out.balance.test[1]$results[,,1][, 1:4]) %>% 
  
  add_row(`treatment=0` = 117, `treatment=1` = 103) %>% 
  
  mutate(condition = c("Liberal", "Republican", "Obama disapprove", 
                  "Ride MBTA every day", "Voted 2010", "Romney voter", "Hispanic threat", 
                   "Age", "Residency year", "College", "Male",  
                   "Hispanic", "White", "Income", "n"))


# Enos just uses xtable(), for which xbal provides a method. But it sure does
# look ugly because xtable() has so few options. I will use gt.

# https://gt.rstudio.com/articles/creating-display-tables.html

# I should do everything with gt going forward. It is obviously the future of
# table making in R. Are they working on regression displays? Still not sure how
# to use gt_preview() interactively.

gt_tbl <- x %>% 
  gt() %>% 
  tab_header(title = "Table 2. Covariate balance across treatment conditions") %>% 
  
  cols_move_to_start(columns = vars(condition)) %>% 
  
  # Set column names. Note that you need to use the original variable names in
  # functions like tab_footnote() even though we call those functions after
  # setting the column names. In other words, these are just for display, they
  # don't effect the data we are passing through the pipe.
  
  cols_label(condition =  "Condition",        
             `treatment=0` = "Control",
             `treatment=1` = "Treatment",
             std.diff = "Standard difference",       
             z = "Z Score") %>% 
  
  # Get the formatting correct.
  
  fmt_number(columns = vars(`treatment=0`, `treatment=1`, std.diff, z),
             decimals = 2) %>% 
  fmt_number(columns = vars(`treatment=0`, `treatment=1`),
             decimals = 0,
             rows = c(14, 15)) %>% 
  fmt_missing(columns = vars(std.diff, z), rows = 15, missing_text = "") %>% 
  
  # Take care of the footnotes. 
  
  tab_options(footnote.glyph = c("*", "&dagger;")) %>%   
  
  tab_footnote(
    footnote = "Difference in standardized units.",
    locations = cells_column_labels(
      columns = vars(std.diff))) %>% 
  
  tab_footnote(
    footnote = "Mean response values for the pretreatment variables accounting for stratification into train stations. All variables are 0 and 1 variables, except for Hispanic threat, which is a seven-point scale indicating how threatening respondents find Hispanics, recoded 0--1; residency, which is measured in years; and income, which is annual income in dollars.",
    locations = cells_data(
      columns = vars(condition),
      rows = 1)) 

gt_tbl %>% as_raw_html() %>% as.character() %>% cat()

```

\newpage

## Figure 2

```{r figure_2}
# I used Gabe Walker's work as a guide in completing this figure.

repeats <- c("numberim","Remain","Englishlan")

x.names <- paste(repeats,".x",sep="")
y.names <- paste(repeats,".y",sep="")

covariates <- c('line.x')

var.names <- c('Number of immigrants be increased?','Children of undocumented be allowed to stay?','English as official language?')

	
##dose response estimates comparison

final.mat <- matrix(nrow = 0, ncol = 8)
subsets <- c('ta', 'tb')


for(subset in subsets){

	out.mat <- matrix(nrow = length(repeats), ncol = 8)
	
	if(subset == 'ta'){
		dat.subset <- dat.all[dat.all$t.time %in% c('t2a', 't4a'),]
	}
	if(subset == 'tb'){
		dat.subset <- dat.all[dat.all$t.time %in% c('t2b', 't4b'),]
	}
	
	z.variable <- 'treatment'
	
	for(j in 1:length(repeats)){
	  
	  # We use this (exact?) same code twice. Clean this up! Or do we need to do it
	  # twice because the universe is different each time and, therefore, the
	  # rescaling? No! It is an absolute, not relative, rescaling.
	  
	  dat.subset$x.new <- (dat.subset[, x.names[j], drop = TRUE]- 1)/4  ##rescale x to 0-1
		dat.subset$y.new <- (dat.subset[, y.names[j], drop = TRUE]- 1)/4  ##rescale y to 0-1
		dat.subset$Y <- dat.subset$y.new - dat.subset$x.new
	
		dat.use <- dat.subset[is.na(dat.subset$Y) == FALSE, ]
				
		x.sd <- sd(dat.use$x.new, na.rm = TRUE)
		x.mean <- mean(dat.use$x.new, na.rm = TRUE)
		
		station.treatment.table <- table(dat.use$station, dat.use[, z.variable, drop = TRUE])
		
		no.control.stations = names(which(station.treatment.table[,1] == 0))
		no.treatment.stations = names(which(station.treatment.table[,2] == 0))
		dat.use = dat.use[!dat.use$station%in%c(no.control.stations,no.treatment.stations),]
		
		dat.use$station = factor(dat.use$station)
		dat.use$treated_unit = factor(dat.use$treated_unit)
		Xs = data.matrix(dat.use[,covariates])
		
		perms <- genperms(Z = dat.use[,z.variable], blockvar=dat.use$station, clustvar=dat.use$treated_unit)
		probs = genprobexact(Z = dat.use[,z.variable], blockvar=dat.use$station, clustvar=dat.use$treated_unit)
		ate = estate(Y = dat.use$Y, Z = dat.use[, z.variable, drop = TRUE], X = Xs, prob = probs)
		Ys = genouts(Y = dat.use$Y, Z = dat.use[, z.variable, drop = TRUE], ate = ate)
		distout <- gendist(Ys,perms, prob=probs)
		disp =	dispdist(distout, ate = ate, display.plot = F)
		
		##fill matrix
		out.mat[j,1] = repeats[j]
		out.mat[j,2] = subset
		out.mat[j,3] = nrow(dat.use)
		out.mat[j,4] = ate
		out.mat[j,5] = x.mean
		out.mat[j,6] = x.sd
		out.mat[j,7] = disp$quantile[1]
		out.mat[j,8] = disp$quantile[2]
		}
		final.mat = rbind(final.mat,out.mat)
	}
final.mat = as.data.frame(final.mat)
colnames(final.mat) = c('variable','subset','N','ate','x.mean','x.sd','quantile.lower','quantile.upper')

final.mat.dose = final.mat ##mat for creating graph later



```



## References
